---
title: "By Channel Across Participants"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  github_document:
    toc: true
  html_document:
    toc: true
params:
  doc_dir: analysis
  data_dir: data/csv_indiv/
  egi_path: data/egi.csv
  harmonic: 1F1
  test_mode: FALSE
  group: Infants
  p_thresh: .01
  plot_titles: FALSE
---

## Description

This document produces scatter plots of the real and imaginary SSVEP components at the harmonic given by the YAML parameter, `params$harmonic`, for all 128 channels. The default harmonic value is `r params$harmonic`.

The document also plots channel-wise results from separate ANOVAs on the real (Sr) and imaginary (Si) components of the SSVEP at the default harmonic.

It also includes barplots of the vector amplitude means (across participants) for results thresholded at $p<$ `r params$p_thresh` as specified in the YAML parameter `params$p_thresh`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
source("R/init_cap.R")
```

```{r Plot-channel-harm}
Plot_channel_harm <- function(chan, harm, df){
  p <- df %>% 
    filter(Channel == chan, Harm == harm) %>% 
    select(iSess, Sr, Si, Pattern, Speed) %>% 
    ggplot() +
    aes(x = Sr, y = Si) +
    geom_point() +
    facet_grid(Speed ~ Pattern) +
    ggtitle(paste0("Channel ", chan, ": ", harm))
  p
}
```

```{r load-merge-clean}
file_list <- (list.files(params$data_dir, pattern = "merged\\.csv$", full.names = TRUE))

# Then run the load_moco() function along this list.
df_moco_list <- lapply(file_list, readr::read_csv)

# Then merge the list of 
df_moco <- Reduce(function(x,y) full_join(x,y, all=TRUE), df_moco_list)

# conds 1-4:
# radial, radial, linear, linear
# slow, fast, slow, fast (5 arcmin/update; 20 arcmin/update; 2deg/s; 8deg/s)

df_moco$Speed <- NA
df_moco$Speed[df_moco$iCond == 1] <- "2deg/s"
df_moco$Speed[df_moco$iCond == 3] <- "2deg/s"
df_moco$Speed[df_moco$iCond == 2] <- "8deg/s"
df_moco$Speed[df_moco$iCond == 4] <- "8deg/s"

df_moco$Pattern <- NA
df_moco$Pattern[df_moco$iCond == 1] <- "Radial"
df_moco$Pattern[df_moco$iCond == 2] <- "Radial"
df_moco$Pattern[df_moco$iCond == 3] <- "Linear"
df_moco$Pattern[df_moco$iCond == 4] <- "Linear"

df_moco$Channel <- as.numeric(substr(df_moco$iCh, 3, 5))

# Drop unused cols
df_moco <- df_moco %>%
  select(iSess, DayAge, Sex, Channel, Harm, Speed, Pattern, Sr, Si)

# Clean-up
rm(df_moco_list, file_list)
```

## Plots

```{r plot-all-channels}
if (params$test_mode) {
  Plot_channel_harm(1, "1F1", df_moco)
} else {
  chans <- 1:128
  lapply(chans, Plot_channel_harm, params$harmonic, df_moco)
}
```

## ANOVA

```{r channel-wise-anova}
#form_maov <- cbind(Sr, Si) ~ Pattern*Speed + Error(iSess)
AOV_by_channel_harm <- function(Ch, H, df, form) {
  df <- df_moco %>%
    filter(Channel == Ch, Harm == H)
  a <- summary(aov(form, df))
  a
}

form_sr <- Sr ~ Pattern*Speed + Error(iSess)
form_si <- Si ~ Pattern*Speed + Error(iSess)

if (params$test_mode) {
  test_1_1F1_Sr <- AOV_by_channel_harm(1,"1F1", df_moco, form_sr)
}
```

### Sr summaries

```{r}
chans <- 1:128
chan_aov_Sr <- lapply(chans, AOV_by_channel_harm, params$harmonic, df_moco, form_sr)
```

### Si summaries

```{r}
chan_aov_Si <- lapply(chans, AOV_by_channel_harm, params$harmonic, df_moco, form_si)
```

### Topo plot

```{r extract-stats}
Extract_stats <- function(model_list) {
  ml = unlist(model_list)
  new_vals <- as.numeric( c( ml['Error: Within.F value1'], 
                             ml['Error: Within.F value2'], 
                             ml['Error: Within.F value3'], 
                             ml['Error: Within.Df1'], 
                             ml['Error: Within.Df2'], 
                             ml['Error: Within.Df3'],
                             ml['Error: Within.Df4'], 
                             ml['Error: Within.Df4'], 
                             ml['Error: Within.Df4'], 
                             ml['Error: Within.Pr(>F)1'], 
                             ml['Error: Within.Pr(>F)2'], 
                             ml['Error: Within.Pr(>F)3']) )
  names(new_vals) <- c("F-Pattern", "F-Speed", "F-Patt*Spd", "df-Pattern", "df-Speed", "df-Patt*Spd", "df-Pattern-denom", "df-Speed-denom", "df-Patt*Spd-denom", "p-Pattern", "p-Speed", "p-Patt*Spd")
  rbind(new_vals)
}

if (params$test_mode) {
  (Extract_stats(test_1_1F1_Sr))
}
```

```{r prep-for-topo-plot}
# Load EGI coordinates
df_egi <- read_csv(params$egi_path)

# Declare function to make statistics data frame from list
Make_stats_df <- function(aov_summ, df_egi, component="Sr") {
  # Extract stats and assemble into data frame
  list_stats <- data.frame(t(sapply(aov_summ, Extract_stats)))
  names(list_stats) <- rep( c("Pattern", "Speed", "Patt_Spd"), 4)
  list_F <- list_stats[,1:3]
  list_dfNum <- list_stats[,4:6]
  list_dfDen <- list_stats[,7:9]
  list_dfp <- list_stats[,10:12]
  
  F_val <- data.frame(list_F) %>% gather(Effect, F_val, Pattern:Patt_Spd)
  dfNum <- data.frame(list_dfNum) %>% gather(Effect, dfNum, Pattern:Patt_Spd)
  dfDen <- data.frame(list_dfDen) %>% gather(Effect, dfDen, Pattern:Patt_Spd)
  pvals <- data.frame(list_dfp) %>% gather(Pvals, pvals, Pattern:Patt_Spd)
  
  df_stats <- data.frame(Chan = rep( 1:128, 3), 
                         Cond = rep(c("Pattern", "Speed", "Patt_Spd"), 
                                    c(128, 128, 128)),
                         Fvals = F_val[,2],
                         FdfNum = dfNum[,2],
                         FdfDen = dfDen[,2],
                         Pvals = pvals[,2],
                         xpos=rep(df_egi$xpos,3),
                         ypos=rep(df_egi$ypos,3))
  
  # Cut pvals
  pvals_cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
  pvals_lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")
  
  # Create cuts based on p-value levels
  Pvals_cuts = cut( df_stats$Pvals, breaks=pvals_cuts, labels=pvals_lbls)
  Pvals_cuts = ordered( Pvals_cuts, levels = rev( pvals_lbls ) )
  df_stats$Pvals_cuts = Pvals_cuts
  
  # Change order of Conditions for plotting
  df_stats$Cond = ordered( df_stats$Cond, levels=c("Pattern", "Speed", "Patt_Spd"))
  
  # Add Sr/Si component
  df_stats$Component <- component
  df_stats
}

if (params$test_mode) {
  (Make_stats_df(chan_aov_Sr, df_egi, "Sr"))
  (Make_stats_df(chan_aov_Si, df_egi, "Si"))
}
```


```{r channel-effects-plot}
Plot_channel_effects <- function(df, harm, group, plot_titles) {
  yquiet = scale_y_continuous("", breaks=NULL)
  xquiet = scale_x_continuous("", breaks=NULL)
  
  # Plot theme for channel topo
  pl_theme_topo <- theme(plot.title = element_text(lineheight=.8, 
                                                   face ="bold", 
                                                   vjust=2, 
                                                   size=rel(1.6)),
                         legend.title=element_text(size=0),
                         legend.text=element_text(size=rel(1)),
                         legend.position="bottom",
                         strip.text = element_text(size = rel(1.5))
                         
  )
  
  pl_title <- paste(init_cap(group), " Group:\nChannel-wise effects for ", harm, sep="" )
  
  # Plot
  pl <- ggplot(data=df, aes(x=xpos, 
                            y=ypos, 
                            color=Pvals_cuts, 
                            size=Pvals_cuts)) +
    geom_point() +
    facet_grid(facets = Component ~ Cond) +
    coord_fixed() + 
    xquiet + 
    yquiet +
    pl_theme_topo
  
  # Add ears, nose
  # topo_ears_nose <- make_ears_nose_overlay(topoplot_fullpath)
  # pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
  
  if (plot_titles) {
    pl + ggtitle(pl_title)
  } else {
    pl
  }
}

if (params$test_mode) {
  (df_chan_stats_Sr <- Make_stats_df(chan_aov_Sr, df_egi, "Sr"))
  (df_chan_stats_Si <- Make_stats_df(chan_aov_Si, df_egi, "Si"))
  (df_stats <- rbind(df_chan_stats_Sr, df_chan_stats_Si))
  Plot_channel_effects(df_stats, "1F1", "Test", TRUE)
} else {
  df_stats <- rbind(Make_stats_df(chan_aov_Sr, df_egi, "Sr"),
                    Make_stats_df(chan_aov_Si, df_egi, "Si"))
  Plot_channel_effects(df_stats, params$this_harm, params$group, params$plot_titles)
}
```

### Mean amplitude plots of `r params$harmonic` channels meeting `r paste('p< ', params$p_thresh, sep='')` for `r init_cap(params$group)` participants.

```{r select-chans-below-thresh-plot}
Select_chans_below <- function(df, cond, p_thresh) {
  df %>% 
    filter(Cond == cond, Pvals < p_thresh) %>%
    select(Chan) %>%
    unique() ->
    chans_below
  as.numeric(unlist(chans_below))
}

Compute_channel_vector_amplitudes <- function(df_chan_stats, df_moco, this_harm, cond, p_thresh) {
  
  chans_below <- Select_chans_below(df_chan_stats, cond, p_thresh)

  if (length(chans_below) > 0) {
  
  # Calculate within participant, then across participant vector means & ampl.
  # This is calculated from full moco dataset.
  # Have to use standard eval commands for dplyr, e.g., group_by_() and quote()
  
  # Within participants...
  df_below_thresh <- df_moco %>%
    filter(Channel %in% chans_below, Harm == this_harm) %>%
    group_by_(quote(Channel), cond, quote(iSess)) %>%
    summarise(sr.sub.mean=mean(Sr), 
              si.sub.mean=mean(Si),
              rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
  # Average across participants  
    group_by_(quote(Channel), cond) %>%
    summarise(sr.group.mean=mean(sr.sub.mean),
              si.group.mean=mean(si.sub.mean),
              sr.group.sd=sd(sr.sub.mean),
              si.group.sd=sd(si.sub.mean),
              rms.amp=sqrt(sr.group.mean^2 + si.group.mean^2),
              rms.amp.sem = mean(rms.amp.sub)/sqrt(n()),
              nsubs=n(),
              sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
              si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  }
}

if (params$test_mode) {
  (Select_chans_below(df_stats, "Speed", .01))
  (Compute_channel_vector_amplitudes(df_stats, df_moco, "1F1", "Speed", .01))
}
```

```{r def-plot-channel-mag-function}
Plot_channel_magnitudes <- function(df_chan_stats, df_moco, 
                                    this_group, this_harm, cond, p_thresh,
                                    plot_titles = FALSE) {
  
  df_plot <- Compute_channel_vector_amplitudes(df_chan_stats, 
                                               df_moco, this_harm, cond,
                                               p_thresh)
  
  if (!(is.null(df_plot))) {
    # Prepare to plot
    limits <- aes(ymax = rms.amp + rms.amp.sem, ymin = rms.amp)
    dodge <- position_dodge(width=0.8)
    pl_title <- paste(init_cap(this_group), 
                      " Group:\nChannels Meeting Criterion for ", 
                      cond, " at ", this_harm, sep="")
    
    pl_theme_bar <- theme(plot.title = element_text(lineheight=.8, 
                                                    face ="bold", 
                                                    vjust=2, size = rel(2.0)),
                          panel.background = element_rect(fill=NA),
                          panel.grid.major = element_blank(),
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(fill=NA, 
                                                      color="black", size=.8,
                                                      linetype="solid"),
                          axis.title.x=element_text(vjust=-.6, size=rel(1.8)),
                          axis.title.y=element_text(face="bold", vjust=1,
                                                    size=rel(1.8)),
                          axis.text=element_text(color="black", size=rel(1.0),
                                                 angle=90),
                          legend.title=element_blank(),
                          legend.text=element_text(size=rel(1.6)), 
                          legend.position="bottom",
                          legend.background=element_blank())
    
    pl <- ggplot(df_plot) +
      aes_string(x="as.factor(Channel)", y="rms.amp", fill=cond) + 
      geom_bar(stat="identity", width=0.8, position=dodge) +
      geom_errorbar(limits, position=dodge, width=0.15) +
      xlab("Channel") +
      ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
      pl_theme_bar
    
    if (plot_titles) {
      pl + ggtitle(pl_title)     
    } else {
      pl
    }
    
  } else {
    cat('No data to plot for ', cond, '.\n', sep="")
    return()
  }
}
```

### Pattern effects: `r params$harmonic`: $p<$ `r params$p_thresh`

```{r figX-vector-amplitude-barplots-pattern}
if (params$test_mode) {
  Plot_channel_magnitudes(df_stats, df_moco, "", "1F1", "Pattern", .01, FALSE)
} else {
  Plot_channel_magnitudes(df_stats, df_moco, params$group, params$harmonic, "Pattern", as.numeric(params$p_thresh), plot_titles=params$plot_titles)
}
```

### Speed effects: `r params$harmonic`: $p<$ `r params$p_thresh`

```{r figX-vector-amplitude-barplots-speed}
if (params$test_mode) {
  Plot_channel_magnitudes(df_stats, df_moco, "", "1F1", "Speed", .01, FALSE)
} else {
  Plot_channel_magnitudes(df_stats, df_moco, params$group,
                          params$harmonic, "Speed",
                          as.numeric(params$p_thresh), 
                          plot_titles=params$plot_titles)
}
```

```{r figX-vector-amplitude-barplots-pattbyspd}
# if (params$test_mode) {
#   Plot_channel_magnitudes(df_stats, df_moco, "", "1F1", "Patt_Spd", .01, FALSE)
# } else {
#   Plot_channel_magnitudes(df_chan_stats, df_moco, params$group,
#                           params$harmonic, "Patt_Spd",
#                           as.numeric(params$p_thresh),
#                           plot_titles=params$plot_titles)
# }
```
